High-level cognition is supported by at least second order dynamic correlations in neural activity patterns

### 2019-09-03 paper/main.tex
backing up progress on results section: decoding results

\begin{figure}[hp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Across-participant decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a. Decoding accuracy as a function of order: PCA.} \textit{Order}
    ($x$-axis) refers to the maximum order of dynamic correlations that were
    available to the classifiers (see \textit{Feature weighting and
      testing}).  The reported across-participant decoding accuracies
    are averaged over all kernel shapes and widths (see
    \textit{Identifying robust decoding results}).  The $y$-values are
    displayed relative to chance accuracy.    The error bars
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upwards
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed test, thresholded at $p <
    0.05$).  The circled values represent the maximum decoding
    accuracy within each experimental condition.
    \textbf{b. Normalized decoding accuracy as a function of order: PCA.}
    This panel displays the same results as Panel a, but here each
    curve has been normalized to have a maximum value of 1 and and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.  Decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d. Noramlized decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.}
  \label{fig:decoding}
\end{figure}

Both sets of temporal decoding analyses yielded qualitatively similar
results for the auditory (non-rest) conditions of the experiment.  The
highest decoding accuracy for participants who listened to the intact
(unscrambled) story was achieved using high-order dynamic correlations (PCA:
second-order; eigenvector-centrality: fourth-order).  Scambled
versions of the story were best decoded by lower-order correlations
(PCA/paragraph: first-order; PCA/word: order zero; eigenvector
centrality/paragraph: order zero; eigenvector centrality/word: order
zero).  The two sets of analyses yielded different results on resting
state data (which we note could be decoded only very slightly above
chance).  The PCA-based approach achieved the highest resting state
decoding accuracy using order zero features (non-correlational
activation-based), whereas the eigenvector centrality-based approach
achieved the highest resting state decoding accuracy using
second-order correlations.

@@

% we used temporal decoders to identify cognitivly relevant neural
% patterns in fMRI data (see forward inference section of methods).
% we explored two approaches to effiently computing high-order
% correlations: PCA (dimensionality reduction) and eigenvector
% centrality (graph measure).  both approaches yielded qualitatively
% similar results for the story listening conditions: intact story was
% best decoded using high-order correlations, whereas scrambled
% stories were best decoded using lower-order correlations.  both
% approaches also yielded slightly above-chance decoding of resting state times
% (perhaps picking up on decreasing attention and/or increasing
% mind-wandering as the resting state scan progressed).  The PCA-based
% approach achieved the highest decoding accuracy using
% non-correlational (activation-based) data, whereas the eigenvector
% centrality approach identified second-order correlations as being
% most informative about resting state times.

>>>>

\begin{figure}[hp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level.pdf}
  \caption{\textbf{Across-participant decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a. Decoding accuracy as a function of order: PCA.}
    \textit{Order} ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    \textit{Feature weighting and testing}).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see \textit{Identifying robust decoding
      results}).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$).  The error bars
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upwards
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed test, thresholded at
    $p < 0.05$).  The circled values represent the maximum decoding
    accuracy within each experimental condition.
    \textbf{b. Normalized decoding accuracy as a function of order:
      PCA.}  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and
    and a minimum value of 0 (including the upper and lower bounds of
    the respective 95\% confidence intervals).  Panels a and b used
    PCA to project each high-dimensional pattern of dynamic
    correlations onto a lower-dimensional space.  \textbf{c.  Decoding
      accuracy as a function of order: eigenvector centrality.} This
    panel is in the same format as Panel a, but here eigenvector
    centrality has been used to project the high-dimensional patterns
    of dynamic correlations onto a lower-dimensional space.
    \textbf{d. Noramlized decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space.}
  \label{fig:decoding}
\end{figure}

Both sets of temporal decoding analyses yielded qualitatively similar
results for the auditory (non-rest) conditions of the experiment
(Fig.~\ref{fig:decoding}: pink, yellow, and teal lines).  The highest
decoding accuracy for participants who listened to the intact
(unscrambled) story was achieved using high-order dynamic correlations
(PCA: second-order; eigenvector-centrality: fourth-order).  Scambled
versions of the story were best decoded by lower-order correlations
(PCA/paragraph: first-order; PCA/word: order zero; eigenvector
centrality/paragraph: order zero; eigenvector centrality/word: order
zero).  The two sets of analyses yielded different decoding results on
resting state data (Fig.~\ref{fig:decoding}: purple lines).  We note
could be decoded only very slightly above chance; we speculate that
the decoders might have picked up on attentional drift, boredom, or
tiredness; we hypothesize these all increase throughought the resting
state scan.  Our decoders might be picking up on aspects of these
loosely defined cognitive states that are common across individuals.
The PCA-based approach achieved the highest resting state decoding
accuracy using order zero features (non-correlational
activation-based), whereas the eigenvector centrality-based approach
achieved the highest resting state decoding accuracy using
second-order correlations.  Taken together, these analyses indicate
that high-level cognitive processing (while listening to the intact
story) is reflected in the dynamics of high-order correlations in
brain activity, whereas lower-level cognitive processing (while
listening to scrambled versions of the story that lack rich meaning)
is reflected in the dynamics of lower-order correlations and
non-correlational activity dynamics.  Further, these patterns are
associated both with the underlying activity patterns (characterized
using PCA) and also with the changing relative positions that
different brain areas occupy in their associated networks
(characterized using eigenvector centrality).

@@
